Cleaning and Processing

### 2024-11-14 code/eda-code/eda-v1.qmd
RRSSM2
Age2; immigrants; year fixed effects

@@

# Recode factor using dplyr
ML_summary$ModifiedColor <- recode_factor(ML_summary$Color, "Parda" = "Afro-Descent", "Preta" = "Afro-Descent", "Branca" = "Euro-Descent")

# Create dummy variables for status and change to English
ML_summary$ModifiedStatus <- recode_factor(ML_summary$Status, "Secundigesta" = "Multiparous", "Trigesta" = "Multiparous", "Multigesta" = "Multiparous", "Secundipara" = "Multiparous", "Multipara" = "Multiparous", "Primigesta" = "Nulliparous", "Nulipara" = "Nulliparous", "Primipara" = "Nulliparous")

# Create dummy variables for nationality and change to English
ML_summary$ModifiedNationality <- recode_factor(ML_summary$Nationality, "Alema" = "European", "Argentina" = "Latin American", "Austriaca" = "European", "Brasileira" = "Brazilian", "Espanhola" = "European", "Francesa" = "European", "Italiana" = "European", "Paraguaya" = "Latin American", "Polaca" = "European", "Portuguesa" = "European", "Rumania" = "European", "Russa" = "European", "Suiça" = "European", "Siria" = "Middle Eastern", "Uruguaya" = "Latin American")

# Create dummy variables for birth weight for infants born to Black, mixed-race, and White mothers for summary table. Create dummy variables for birth weight according to fetal sex.
ML_summary <- ML_summary %>%
  mutate(
    Weightblack = factor(if_else(Color == "Preta", "Yes", "No"), levels = c("No", "Yes")),
    Weightmixed = factor(if_else(Color == "Parda", "Yes", "No"), levels = c("No", "Yes")),
    Weightwhite = factor(if_else(Color == "Branca", "Yes", "No"), levels = c("No", "Yes")),
    Weightfemale = factor(if_else(Sex == "F", "Yes", "No"), levels = c("No", "Yes")),
    Weightmale = factor(if_else(Sex == "M", "Yes", "No"), levels = c("No", "Yes")),
    WeightgramsBlack = if_else(Weightblack == "Yes", Weightgrams, NA_real_),
    WeightgramsWhite = if_else(Weightwhite == "Yes", Weightgrams, NA_real_),
    WeightgramsMixed = if_else(Weightmixed == "Yes", Weightgrams, NA_real_),
    WeightgramsFemale = if_else(Weightfemale == "Yes", Weightgrams, NA_real_),
    WeightgramsMale = if_else(Weightmale == "Yes", Weightgrams, NA_real_),
    WeightgramsAfro = if_else(ModifiedColor == "Afro-Descent", Weightgrams, NA_real_),
    WeightgramsEuro = if_else(ModifiedColor == "Euro-Descent", Weightgrams, NA_real_)
  )

@@

# View the updated dataframe
print(head(ML_summary))
```
I will now change the observations from Portuguese to English for the remaining observations to clean up things for the summary table. In the future, I should perform this step in the processing script.

```{r}
# Change rest of observations from Portuguese to English
ML_summary <- ML_summary %>%
  mutate(
    Color = recode_factor(Color, "Preta" = "Black", "Parda" = "Mixed Race", "Branca" = "White"),
    Nationality = recode_factor(Nationality, "Alema" = "German", "Argentina" = "Argentine", "Austriaca" = "Austrian", "Brasileira" = "Brazilian", "Espanhola" = "Spanish", "Francesa" = "French", "Italiana" = "Italian", "Paraguaya" = "Paraguayan", "Polaca" = "Polish", "Portuguesa" = "Portuguese", "Rumania" = "Romanian", "Russa" = "Russian", "Suiça" = "Swiss", "Siria" = "Syrian", "Uruguaya" = "Uruguayan"),
    Birth = recode_factor(Birth, "aborto" = "Abortion", "intervencionista" = "Interventionist", "natural" = "Natural", "operatório" = "Operative"),
    MaternalOutcome = recode_factor(MaternalOutcome, "alta" = "Discharged", "morta" = "Death", "transferência" = "Hospital transferal"),
    FetalOutcome = recode_factor(FetalOutcome, "vivo" = "Live Birth", "morto" = "Stillbirth"),
    Status = recode_factor(Status, "Multigesta" = "Multigravida", "Trigesta" = "Trigravida", "Secundigesta" = "Secundigravida", "Secundipara" = "Secundiparous", "Multipara" = "Multiparous", "Primigesta" = "Primigravida", "Nulipara" = "Nulliparous", "Primipara" = "Primiparous")
    )

# Summarize missing data using dplyr
missing_data_summary <- ML_summary %>% dplyr::summarise_all(list(~sum(is.na(.))))
missing_data_summary

@@

`Gtsummary` is a very easy way to create summary statistics (Table 1). For previous code I worked on in relation to summary stats, see the `edacode-extra.R` file.

```{r, output="html"}
# Create gtsummary table full table for supplemental analysis
table_supplemental <- 
  ML_summary %>%
  tbl_summary(
    include = c(Color, ModifiedColor, Status, ModifiedStatus, Age, Nationality, ModifiedNationality, Birth, MaternalOutcome, FetalOutcome, Sex, Lengthcentimeters, Weightgrams),
    missing = "no",
    label = list(ModifiedColor ~ "Ancestry", Status ~ "Parity or Gravidity", ModifiedStatus ~ "Parity", Age ~ "Maternal Age", ModifiedNationality ~ "Combined Nationality", Birth ~ "Birth Outcome", MaternalOutcome ~ "Maternal Outcome", FetalOutcome ~ "Fetal Outcome", Lengthcentimeters ~ "Birth Length (cms)", Weightgrams ~ "Birth Weight (grms)"),
    statistic = list(all_continuous() ~ "{mean} ({p25}, {p75})")
  ) %>%
  add_n()
table_supplemental

@@

# Create gtsummary table for final analysis
table1 <- 
  ML_summary %>%
  mutate(
    WeightgramsBlack = if_else(Weightblack == "Yes", Weightgrams, NA_real_),
    WeightgramsWhite = if_else(Weightwhite == "Yes", Weightgrams, NA_real_),
    WeightgramsMixed = if_else(Weightmixed == "Yes", Weightgrams, NA_real_),
    WeightgramsAfro = if_else(ModifiedColor == "Afro-Descent", Weightgrams, NA_real_),
    WeightgramsEuro = if_else(ModifiedColor == "Euro-Descent", Weightgrams, NA_real_),
    WeightgramsFemale = if_else(Weightfemale == "Yes", Weightgrams, NA_real_),
    WeightgramsMale = if_else(Weightmale == "Yes", Weightgrams, NA_real_)
  ) %>%
  tbl_summary(
    include = c(Color, ModifiedColor, ModifiedStatus, Age, ModifiedNationality, Birth, MaternalOutcome, FetalOutcome, Sex, Lengthcentimeters, Weightgrams, WeightgramsBlack, WeightgramsMixed, WeightgramsWhite, WeightgramsAfro, WeightgramsEuro, WeightgramsFemale, WeightgramsMale, BirthweightCategory),
    missing = "no",
    label = list(
      Color ~ "Color",
      ModifiedColor ~ "Ancestry", 
      ModifiedStatus ~ "Parity", 
      Age ~ "Maternal Age", 
      ModifiedNationality ~ "Combined Nationality", 
      Birth ~ "Birth Outcome", 
      MaternalOutcome ~ "Maternal Outcome", 
      FetalOutcome ~ "Fetal Outcome", 
      Sex ~ "Sex",
      Lengthcentimeters ~ "Birth Length (cm)", 
      Weightgrams ~ "Birth Weight (g)",
      WeightgramsBlack ~ "Black",
      WeightgramsMixed ~ "Mixed-Race",
      WeightgramsWhite ~ "White",
      WeightgramsAfro ~ "Afro-Descent",
      WeightgramsEuro ~ "Euro-Descent",
      WeightgramsFemale ~ "Female",
      WeightgramsMale ~ "Male",
      BirthweightCategory ~ "Birth Weight Category"
    ),
    statistic = list(
       all_continuous() ~ "{mean} ({p25}, {p75})",
      WeightgramsBlack ~ "{mean} ({p25}, {p75})",
      WeightgramsMixed ~ "{mean} ({p25}, {p75})",
      WeightgramsWhite ~ "{mean} ({p25}, {p75})",
      WeightgramsFemale ~ "{mean} ({p25}, {p75})",
      WeightgramsMale ~ "{mean} ({p25}, {p75})"
    ),
    digits = list(WeightgramsBlack ~ 0, WeightgramsMixed ~ 0, WeightgramsWhite ~ 0, WeightgramsFemale ~ 0, WeightgramsMale ~ 0)
  ) %>%
  add_n() %>%
  modify_table_body( #indent these three rows
    ~.x %>%
      mutate(label = case_when(
        variable == "WeightgramsBlack" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsWhite" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsMixed" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsAfro" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsEuro" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsFemale" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsMale" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        TRUE ~ label
      ))
  )
table1

@@

```{r}
# Calculate frequencies of birth category without abortion
birth_freq2 <- ML_summary %>%
  filter(!is.na(Birth)) %>% # Exclude rows with NA in birth variable
  filter(Birth != "abortion") %>% # Exclude rows with abortion
  count(Birth, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(birth_freq2)
```

@@

>>>>

I will now change the observations from Portuguese to English for the remaining observations to clean up things for the summary table. In the future, I should perform this step in the processing script.

@@

# Change observations from Portuguese to English
ML_summary <- ML_summary %>%
  mutate(
    Color = recode_factor(Color, "Preta" = "Black", "Parda" = "Mixed Race", "Branca" = "White"),
    Nationality = recode_factor(Nationality, "Alema" = "German", "Argentina" = "Argentine", "Austriaca" = "Austrian", "Brasileira" = "Brazilian", "Espanhola" = "Spanish", "Francesa" = "French", "Italiana" = "Italian", "Paraguaya" = "Paraguayan", "Polaca" = "Polish", "Portuguesa" = "Portuguese", "Rumania" = "Romanian", "Russa" = "Russian", "Suiça" = "Swiss", "Siria" = "Syrian", "Uruguaya" = "Uruguayan"),
    Birth = recode_factor(Birth, "aborto" = "Abortion", "intervencionista" = "Interventionist", "natural" = "Natural", "operatório" = "Operative"),
    MaternalOutcome = recode_factor(MaternalOutcome, "alta" = "Discharged", "morta" = "Death", "transferência" = "Hospital transferal"),
    FetalOutcome = recode_factor(FetalOutcome, "vivo" = "Live Birth", "morto" = "Stillbirth"),
    Status = recode_factor(Status, "Multigesta" = "Multigravida", "Trigesta" = "Trigravida", "Secundigesta" = "Secundigravida", "Secundipara" = "Secundiparous", "Multipara" = "Multiparous", "Primigesta" = "Primigravida", "Nulipara" = "Nulliparous", "Primipara" = "Primiparous")
    )

# Recode factor using dplyr
ML_summary$ModifiedColor <- recode_factor(ML_summary$Color, "Mixed Race" = "Afro-Descent", "Black" = "Afro-Descent", "White" = "Euro-Descent")

# Create dummy variables for status and change to English
ML_summary$ModifiedStatus <- recode_factor(ML_summary$Status, "Secundigravida" = "Multiparous", "Trigravida" = "Multiparous", "Multigravida" = "Multiparous", "Secundiparous" = "Multiparous", "Multiparous" = "Multiparous", "Primigravida" = "Nulliparous", "Nulliparous" = "Nulliparous", "Primiparous" = "Nulliparous")

# Create dummy variables for nationality
ML_summary$ModifiedNationality <- recode_factor(ML_summary$Nationality, "German" = "European", "Argentine" = "Latin American", "Austrian" = "European", "Brazilian" = "Brazilian", "Spanish" = "European", "French" = "European", "Italian" = "European", "Paraguayan" = "Latin American", "Polish" = "European", "Portuguese" = "European", "Romanian" = "European", "Russian" = "European", "Swiss" = "European", "Syrian" = "Middle Eastern", "Uruguayan" = "Latin American")

# Create dummy variables for birth weight for infants born to Black, mixed-race, and White mothers for summary table. Create dummy variables for birth weight according to fetal sex.
ML_summary <- ML_summary %>%
  mutate(
    Weightblack = factor(if_else(Color == "Black", "Yes", "No"), levels = c("No", "Yes")),
    Weightmixed = factor(if_else(Color == "Mixed Race", "Yes", "No"), levels = c("No", "Yes")),
    Weightwhite = factor(if_else(Color == "White", "Yes", "No"), levels = c("No", "Yes")),
    Weightfemale = factor(if_else(Sex == "F", "Yes", "No"), levels = c("No", "Yes")),
    Weightmale = factor(if_else(Sex == "M", "Yes", "No"), levels = c("No", "Yes")),
    WeightgramsBlack = if_else(Weightblack == "Yes", Weightgrams, NA_real_),
    WeightgramsWhite = if_else(Weightwhite == "Yes", Weightgrams, NA_real_),
    WeightgramsMixed = if_else(Weightmixed == "Yes", Weightgrams, NA_real_),
    WeightgramsFemale = if_else(Weightfemale == "Yes", Weightgrams, NA_real_),
    WeightgramsMale = if_else(Weightmale == "Yes", Weightgrams, NA_real_),
    WeightgramsAfro = if_else(ModifiedColor == "Afro-Descent", Weightgrams, NA_real_),
    WeightgramsEuro = if_else(ModifiedColor == "Euro-Descent", Weightgrams, NA_real_)
  )

@@

# Create dummy variables for birth weight for infants born to mothers of different nationalities for summary table.
ML_summary <- ML_summary %>%
  mutate( # First create yes/not dummy variables for count
    WeightPor = factor(if_else(Nationality == "Portuguese", "Yes", "No"), levels = c("No", "Yes")),
    WeightGer = factor(if_else(Nationality == "German", "Yes", "No"), levels = c("No", "Yes")),
    WeightIta = factor(if_else(Nationality == "Italian", "Yes", "No"), levels = c("No", "Yes")),
    WeightAus = factor(if_else(Nationality == "Austrian", "Yes", "No"), levels = c("No", "Yes")),
    WeightFr = factor(if_else(Nationality == "French", "Yes", "No"), levels = c("No", "Yes")),
    WeightSp = factor(if_else(Nationality == "Spanish", "Yes", "No"), levels = c("No", "Yes")),
    WeightPol = factor(if_else(Nationality == "Polish", "Yes", "No"), levels = c("No", "Yes")),
    WeightSw = factor(if_else(Nationality == "Swiss", "Yes", "No"), levels = c("No", "Yes")),
    WeightRom = factor(if_else(Nationality == "Romanian", "Yes", "No"), levels = c("No", "Yes")),
    WeightRus = factor(if_else(Nationality == "Russian", "Yes", "No"), levels = c("No", "Yes")),
    WeightArg = factor(if_else(Nationality == "Argentine", "Yes", "No"), levels = c("No", "Yes")),
    WeightPar = factor(if_else(Nationality == "Paraguayan", "Yes", "No"), levels = c("No", "Yes")),
    WeightUr = factor(if_else(Nationality == "Uruguayan", "Yes", "No"), levels = c("No", "Yes")),
    WeightSy = factor(if_else(Nationality == "Syrian", "Yes", "No"), levels = c("No", "Yes")),
    WeightEuro = factor(if_else(Nationality == "German" | 
                                  Nationality == "Italian" |
                                  Nationality == "Austrian" |
                                  Nationality == "French" |
                                  Nationality == "Spanish" |
                                  Nationality == "Polish" |
                                  Nationality == "Swiss" |
                                  Nationality == "Romanian" |
                                  Nationality == "Russian", "Yes", "No"), levels = c("No", "Yes")),
    WeightBrWh = factor(if_else(Nationality == "Brazilian" & Color == "White", "Yes", "No"), levels = c("No", "Yes")),
    WeightBrBl = factor(if_else(Nationality == "Brazilian" & Color == "Black", "Yes", "No"), levels = c("No", "Yes")),
    WeightBrMx = factor(if_else(Nationality == "Brazilian" & Color == "Mixed Race", "Yes", "No"), levels = c("No", "Yes")),
    WeightgramsPor = if_else(Nationality == "Portuguese", Weightgrams, NA_real_), # then create weight variables
    WeightgramsGer = if_else(Nationality == "German", Weightgrams, NA_real_),
    WeightgramsIta = if_else(Nationality == "Italian", Weightgrams, NA_real_),
    WeightgramsAus = if_else(Nationality == "Austrian", Weightgrams, NA_real_),
    WeightgramsFr = if_else(Nationality == "French", Weightgrams, NA_real_),
    WeightgramsSp = if_else(Nationality == "Spanish", Weightgrams, NA_real_),
    WeightgramsPol = if_else(Nationality == "Polish", Weightgrams, NA_real_),
    WeightgramsSw = if_else(Nationality == "Swiss", Weightgrams, NA_real_),
    WeightgramsRom = if_else(Nationality == "Romanian", Weightgrams, NA_real_),
    WeightgramsRus = if_else(Nationality == "Russian", Weightgrams, NA_real_),
    WeightgramsArg = if_else(Nationality == "Argentine", Weightgrams, NA_real_),
    WeightgramsPar = if_else(Nationality == "Paraguayan", Weightgrams, NA_real_),
    WeightgramsUr = if_else(Nationality == "Uruguayan", Weightgrams, NA_real_),
    WeightgramsSy = if_else(Nationality == "Syrian", Weightgrams, NA_real_),
    WeightgramsEuro = if_else(WeightEuro == "Yes", Weightgrams, NA_real_),
    WeightgramsBrWh = if_else(Nationality == "Brazilian" & Color == "White", Weightgrams, NA_real_),
    WeightgramsBrBl = if_else(Nationality == "Brazilian" & Color == "Black", Weightgrams, NA_real_),
    WeightgramsBrMx = if_else(Nationality == "Brazilian" & Color == "Mixed Race", Weightgrams, NA_real_)
  )

# Verify new variables
str(ML_summary[, c("WeightPor", "WeightgramsPor", 
                   "WeightGer", "WeightgramsGer", 
                   "WeightIta", "WeightgramsIta", 
                   "WeightAus", "WeightgramsAus", 
                   "WeightFr", "WeightgramsFr", 
                   "WeightSp", "WeightgramsSp", 
                   "WeightPol", "WeightgramsPol", 
                   "WeightSw", "WeightgramsSw",
                   "WeightRom", "WeightgramsRom",
                   "WeightRus", "WeightgramsRus",
                   "WeightArg", "WeightgramsArg",
                   "WeightPar", "WeightgramsPar",
                   "WeightUr", "WeightgramsUr",
                   "WeightSy", "WeightgramsSy",
                   "WeightEuro", "WeightgramsEuro",
                   "WeightBrWh", "WeightgramsBrWh", 
                   "WeightBrBl", "WeightgramsBrBl", 
                   "WeightBrMx", "WeightgramsBrMx")])

# View the updated dataframe
print(head(ML_summary))

# Calculate both mean and median weights by nationality
Euro_stats <- ML_summary %>%
  reframe(
    Portuguese = list(
      n = sum(!is.na(WeightgramsPor)),
      mean = mean(WeightgramsPor, na.rm = TRUE),
      median = median(WeightgramsPor, na.rm = TRUE)
    ),
    German = list(
      n = sum(!is.na(WeightgramsGer)),
      mean = mean(WeightgramsGer, na.rm = TRUE),
      median = median(WeightgramsGer, na.rm = TRUE)
    ),
    Italian = list(
      n = sum(!is.na(WeightgramsIta)),
      mean = mean(WeightgramsIta, na.rm = TRUE),
      median = median(WeightgramsIta, na.rm = TRUE)
    ),
    Austrian = list(
      n = sum(!is.na(WeightgramsAus)),
      mean = mean(WeightgramsAus, na.rm = TRUE),
      median = median(WeightgramsAus, na.rm = TRUE)
    ),
    French = list(
      n = sum(!is.na(WeightgramsFr)),
      mean = mean(WeightgramsFr, na.rm = TRUE),
      median = median(WeightgramsFr, na.rm = TRUE)
    ),
    Spanish = list(
      n = sum(!is.na(WeightgramsSp)),
      mean = mean(WeightgramsSp, na.rm = TRUE),
      median = median(WeightgramsSp, na.rm = TRUE)
    ),
    Polish = list(
      n = sum(!is.na(WeightgramsPol)),
      mean = mean(WeightgramsPol, na.rm = TRUE),
      median = median(WeightgramsPol, na.rm = TRUE)
    ),
    Swiss = list(
      n = sum(!is.na(WeightgramsSw)),
      mean = mean(WeightgramsSw, na.rm = TRUE),
      median = median(WeightgramsSw, na.rm = TRUE)
    ),
    Romanian = list(
      n = sum(!is.na(WeightgramsRom)),
      mean = mean(WeightgramsRom, na.rm = TRUE),
      median = median(WeightgramsRom, na.rm = TRUE)
    ),
    Russian = list(
      n = sum(!is.na(WeightgramsRus)),
      mean = mean(WeightgramsRus, na.rm = TRUE),
      median = median(WeightgramsRus, na.rm = TRUE)
    ),
    Euro = list(
      n = sum(!is.na(WeightgramsEuro)),
      mean = mean(WeightgramsEuro, na.rm = TRUE),
      median = median(WeightgramsEuro, na.rm = TRUE)
    ),
    Argentine = list(
      n = sum(!is.na(WeightgramsArg)),
      mean = mean(WeightgramsArg, na.rm = TRUE),
      median = median(WeightgramsArg, na.rm = TRUE)
    ),
    Paraguayan = list(
      n = sum(!is.na(WeightgramsPar)),
      mean = mean(WeightgramsPar, na.rm = TRUE),
      median = median(WeightgramsPar, na.rm = TRUE)
    ),
    Uruguayan = list(
      n = sum(!is.na(WeightgramsUr)),
      mean = mean(WeightgramsUr, na.rm = TRUE),
      median = median(WeightgramsUr, na.rm = TRUE)
    ),
    Syrian = list(
      n = sum(!is.na(WeightgramsSy)),
      mean = mean(WeightgramsSy, na.rm = TRUE),
      median = median(WeightgramsSy, na.rm = TRUE)
    ),
    Brazil_white = list(
      n = sum(!is.na(WeightgramsBrWh)),
      mean = mean(WeightgramsBrWh, na.rm = TRUE),
      median = median(WeightgramsBrWh, na.rm = TRUE)
    ),
    Brazil_black = list(
      n = sum(!is.na(WeightgramsBrBl)),
      mean = mean(WeightgramsBrBl, na.rm = TRUE),
      median = median(WeightgramsBrBl, na.rm = TRUE)
    ),
    Brazil_mixed = list(
      n = sum(!is.na(WeightgramsBrMx)),
      mean = mean(WeightgramsBrMx, na.rm = TRUE),
      median = median(WeightgramsBrMx, na.rm = TRUE)
    )
  ) %>%
  tidyr::unnest_longer(col = everything())
Euro_stats

# T-test between Portuguese and all other European nationalities combined
t_test_euro <- t.test(ML_summary$WeightgramsPor, ML_summary$WeightgramsEuro)
t_test_euro

t_test_brazil <- t.test(ML_summary$WeightgramsPor, ML_summary$WeightgramsBrWh)
t_test_brazil

# Put t-test into gtsummary table for supplemental analysis
# Create a table with both t-tests
table_ttest <- tibble(
  Comparison = c("Portuguese vs Other European", "Portuguese vs Brazilian White"),
  `t-statistic` = c(t_test_euro$statistic, t_test_brazil$statistic),
  `p-value` = c(t_test_euro$p.value, t_test_brazil$p.value),
  `Mean Difference` = c(diff(t_test_euro$estimate), diff(t_test_brazil$estimate)),
  `95% CI Lower` = c(t_test_euro$conf.int[1], t_test_brazil$conf.int[1]),
  `95% CI Upper` = c(t_test_euro$conf.int[2], t_test_brazil$conf.int[2])
    ) %>%
  gt::gt() %>%
  gt::fmt_number(
    columns = c(`t-statistic`),
    decimals = 2
  ) %>%
  gt::fmt_number(
    columns = c(`p-value`),
    decimals = 3
  ) %>%
  gt::fmt_number(
    columns = c(`Mean Difference`, `95% CI Lower`, `95% CI Upper`),
    decimals = 1
  ) %>%
  gt::tab_header(
    title = "T-Test Results"
  )
table_ttest

# Save the table
saveRDS(table_ttest, file = here::here("results", "tables", "table_ttest_final.rds"))

# Put means into gtsummary table for supplemental analysis
table_nationality_stats <- ML_summary %>%
  select(WeightgramsPor, WeightgramsGer, WeightgramsIta, WeightgramsAus,
         WeightgramsFr, WeightgramsSp, WeightgramsPol, WeightgramsSw,
         WeightgramsRom, WeightgramsRus, WeightgramsEuro, WeightgramsArg,
         WeightgramsPar, WeightgramsUr, WeightgramsSy, WeightgramsBrWh,
         WeightgramsBrBl, WeightgramsBrMx) %>%
  tbl_summary(
    missing = "no",
    label = list(
      WeightgramsPor ~ "Portuguese",
      WeightgramsGer ~ "German",
      WeightgramsIta ~ "Italian",
      WeightgramsAus ~ "Austrian",
      WeightgramsFr ~ "French",
      WeightgramsSp ~ "Spanish",
      WeightgramsPol ~ "Polish",
      WeightgramsSw ~ "Swiss",
      WeightgramsRom ~ "Romanian",
      WeightgramsRus ~ "Russian",
      WeightgramsEuro ~ "European (Exc. Portuguese)",
      WeightgramsArg ~ "Argentine",
      WeightgramsPar ~ "Paraguayan",
      WeightgramsUr ~ "Uruguayan",
      WeightgramsSy ~ "Syrian",
      WeightgramsBrWh ~ "Brazilian White",
      WeightgramsBrBl ~ "Brazilian Black",
      WeightgramsBrMx ~ "Brazilian Mixed-Race"
    ),
    type = list(everything() ~ "continuous"),
    statistic = list(all_continuous() ~ "{mean} ({p25}, {p75})"),
    digits = everything() ~ 0 # all numbers to 0 decimal places
  ) %>%
  add_n() %>%
  modify_header(label ~ "**Nationality**",
                stat_0 ~ "**Mean (Q1, Q3)**",
                n ~ "**N**")
table_nationality_stats

# Save table for supplemental analysis
saveRDS(table_nationality_stats, file = here::here("results", "tables", "table_nationality_stats_final.rds"))
```
The t-tests demonstrate no significant difference in the mean birthweight between Portuguese and all other European immigrants and between Portuguese and White Brazilian women. This suggests that differentiating European nationality further will not influence the regression results.

Further, because there is only 1 observation for Austrian nationality; 3 observations for French nationality; and 2 observations for Polish and Swiss nationalities, I will drop these from my larger summary table.

```{r}
# Summarize missing data using dplyr
missing_data_summary <- ML_summary %>% dplyr::summarise_all(list(~sum(is.na(.))))
missing_data_summary

@@

`Gtsummary` is a very easy way to create summary statistics (Table 1). For previous code I worked on in relation to summary stats, see the `edacode-extra.R` file. I have excluded all nationalities with sample sizes smaller than 10 for infant birth weight (Argentine, Paraguayan, Syrian, Uruguayan, Romanian, French, Polish, Swiss, and Austrian).

```{r, output="html"}
# Create gtsummary table full table for supplemental analysis
table_supplemental <- 
  ML_summary %>%
  tbl_summary(
    include = c(Color, Status, Age, Nationality, Birth, MaternalOutcome, FetalOutcome, Sex, Lengthcentimeters, Weightgrams),
    missing = "no",
    label = list(
      Color ~ "Skin color",
      Status ~ "Parity or Gravidity",
      Age ~ "Maternal Age",
      Nationality ~ "Nationality",
      Birth ~ "Birth Outcome", 
      MaternalOutcome ~ "Maternal Outcome", 
      FetalOutcome ~ "Fetal Outcome",
      Sex ~ "Infant Sex",
      Lengthcentimeters ~ "Birth Length (cms)", 
      Weightgrams ~ "Birth Weight (grms)"
      ),
    statistic = list(all_continuous() ~ "{mean} ({p25}, {p75})")
  ) %>%
  add_n()
table_supplemental

@@

# Create gtsummary table for final analysis
table1 <- 
  ML_summary %>%
  tbl_summary(
    include = c(Color, ModifiedColor, ModifiedStatus, Age, ModifiedNationality, Birth, MaternalOutcome, FetalOutcome, Sex, Lengthcentimeters, Weightgrams, WeightgramsBlack, WeightgramsMixed, WeightgramsWhite, WeightgramsAfro, WeightgramsEuro, WeightgramsPor, WeightgramsGer, WeightgramsIta, WeightgramsSp, WeightgramsRus, WeightgramsBrWh, WeightgramsFemale, WeightgramsMale, BirthweightCategory),
    missing = "no",
    type = list(WeightgramsRus ~ "continuous"),
    label = list(
      Color ~ "Color",
      ModifiedColor ~ "Ancestry", 
      ModifiedStatus ~ "Parity", 
      Age ~ "Maternal Age", 
      ModifiedNationality ~ "Combined Nationality", 
      Birth ~ "Birth Outcome", 
      MaternalOutcome ~ "Maternal Outcome", 
      FetalOutcome ~ "Fetal Outcome", 
      Sex ~ "Sex",
      Lengthcentimeters ~ "Birth Length (cm)", 
      Weightgrams ~ "Birth Weight (g)",
      WeightgramsBlack ~ "Black",
      WeightgramsMixed ~ "Mixed-Race",
      WeightgramsWhite ~ "White",
      WeightgramsAfro ~ "Afro-Descent",
      WeightgramsEuro ~ "Euro-Descent",
      WeightgramsPor ~ "Portuguese",
      WeightgramsGer ~ "German",
      WeightgramsIta ~ "Italian",
      WeightgramsSp ~ "Spanish",
      WeightgramsRus ~ "Russian",
      WeightgramsBrWh ~ "White Brazilian",
      WeightgramsFemale ~ "Female",
      WeightgramsMale ~ "Male",
      BirthweightCategory ~ "Birth Weight Category"
    ),
    statistic = list(
       all_continuous() ~ "{mean} ({p25}, {p75})",
       WeightgramsRus ~ "{mean} ({p25}, {p75})"
    )
  ) %>%
  add_n() %>%
  modify_table_body( #indent these three rows
    ~.x %>%
      mutate(label = case_when(
        variable == "WeightgramsBlack" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsWhite" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsMixed" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsAfro" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsEuro" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsPor" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsGer" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsIta" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsSp" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsRus" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsBrWh" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsFemale" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        variable == "WeightgramsMale" ~ paste0("\u00A0\u00A0\u00A0\u00A0", label),
        TRUE ~ label
      )) 
    )
table1

@@

```{r}
# Calculate frequencies of birth category without abortion
birth_freq2 <- ML_summary %>%
  filter(!is.na(Birth)) %>% # Exclude rows with NA in birth variable
  filter(Birth != "Abortion") %>% # Exclude rows with abortion
  count(Birth, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(birth_freq2)
```

@@

Now I will compute total fertility rates for Rio de Janeiro, São Paulo, and Brazil from 1940 data.

```{r}
# Path to summary data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","Fertility_Rates.rds")

# Load data
Fertility_Rates <- readRDS(data_location)

# Check structure
str(Fertility_Rates)

tfr_region <- Fertility_Rates %>%
  group_by(Region) %>%
  summarize(TFR = round(sum(`Age-Specific Rate`)/100, 1))

# To view the results
print(tfr_region)
```

### 2024-11-14 code/processing-code/processingfile-v1.qmd
RRSSM2
Age2; immigrants; year fixed effects

# Save Mean_Birthweight as RDS
save_data_location <- here::here("data","processed-data","Mean_Birthweights.rds")
saveRDS(Mean_Birthweights, file = save_data_location)
```
# Notes

>>>>

# Save Mean_Birthweight as RDS
save_data_location <- here::here("data","processed-data","Mean_Birthweights.rds")
saveRDS(Mean_Birthweights, file = save_data_location)
```
# Data loading for total fertility rates (TFR)

I will also discuss total fertility rates (TFR) for Rio de Janeiro and Brazil.

```{r}
# Path to data
# Note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","1940_FertilityRates.csv")
Fertility_Rates <- read_csv(data_location)

# Structure of Data
glimpse(Fertility_Rates)

# Save Mean_Birthweight as RDS
save_data_location <- here::here("data","processed-data","Fertility_Rates.rds")
saveRDS(Fertility_Rates, file = save_data_location)
```
